# Generating and Summarizing Data

For this process we have 2 new treatments, when taken a patient will see improvement within 15 days. Our company wants to find which one is more likely to see improvement within the first 5 days. Using the given [techniques](#sec-tech), we want to communicate which one is more likely within the given timeframe. To do this we will be analyzing a simulated data set on the treatments, which contains participants in each treatment group(Trt) and how long it takes for them to see improvements(Days)

-   First, we will need to read in the data set. This will be used with our future functions, but can be substituted by another data set. This data set uses two discrete variables, but you are able to use one discrete and one continuous as well. 

```{r}
trtdata<- read.csv("http://raw.githubusercontent.com/unl-statistics/annual-review-user-guide-Nathan-Swenson/main/trtdata.csv")
```

- Second, we will need to find the proportion in which each treatment sees improvement in the first 5 days, and then we will find the difference of the 2 proportions. This will give us the more effective treatment.

```{r}

Trt2_proportion<- sum(
  trtdata$Days[trtdata$Trt == "Trt2"] >= 1 & 
    trtdata$Days[trtdata$Trt == "Trt2"] <= 5)/
  length(trtdata$Days[trtdata$Trt == "Trt2"])

Trt1_proportion<- sum(
  trtdata$Days[trtdata$Trt == "Trt1"] >= 1 & 
    trtdata$Days[trtdata$Trt == "Trt1"] <= 5)/
  length(trtdata$Days[trtdata$Trt == "Trt1"])

difference<- Trt1_proportion - Trt2_proportion

print(difference)
```

- Third, we would like to calculate the p-value. This will allow us to determine whether there is a significant difference of proportions. To do this we will run a chi-squared test and then index the p_value. (where x is the treatment within the first 5 days and n is number of observations for each)

```{r}
trt_test<- prop.test(
  x=c(
    sum(trtdata$Days[trtdata$Trt == "Trt1"] >= 1 & 
          trtdata$Days[trtdata$Trt == "Trt1"] <= 5), 
    sum(trtdata$Days[trtdata$Trt == "Trt2"] >= 1 & 
          trtdata$Days[trtdata$Trt == "Trt2"] <= 5)), 
  n = c(500, 500))

print(p_value<- trt_test$p.value)
```