# Creating Graphics from the Summarized Data {#sec-graph}

```{r, echo = FALSE}
trtdata<- read.csv("trtdata.csv")
```

Graphics can greatly improve comprehension of risk and can be very effective when paired with verbal communication. This section will help you visualize a graph that can complement your communicated risk.

- First, you will need to install and load the following packages. These packages will be used in data manipulation and the plot generation.

```{r, eval = FALSE, echo=FALSE}
install.packages("distributional")

install.packages("dplyr")

install.packages("ggdibbler", dep=TRUE)

install.packages("ggplot2")
```

- Second, you will need to get the distributional variables, so we can graph the variability in the graph. To do this we will create a table of the proportions so we can create the distribute for each treatment type.

```{r, echo=FALSE}

library(dplyr)
library(distributional)

trtdata<- trtdata |> group_by(Trt)
prop<- as.data.frame(prop.table(table(trtdata[2:3])))


Trt2_dist <- dist_categorical(
  list(c(prop$Freq[prop$Trt == "Trt2"]))
)
Trt1_dist <- dist_categorical(
  list(c(prop$Freq[prop$Trt == "Trt1"]))
)

trtdata <- trtdata |>
  mutate(
    Day_dist = if_else(Trt == "Trt2",
                       Trt2_dist,
                       Trt1_dist)
  )
```


- Finally, we will be creating our graph. We will be creating a bar chart, mapping the treatments to orange and blue. The fading represents the variance of how many participants see improvements on the given day.

```{r}
library("ggdibbler")
library("ggplot2")

ggplot(trtdata, aes(x=Day_dist)) +
  geom_bar_sample(
    aes(fill = Trt),
    alpha = 0.1,
    position = "dodge_identity") +
    scale_fill_manual(values = c("Trt2" = "#0072B2", 
                                 "Trt1" = "#E69F00")) +
  ggtitle("Improvements Occur over Days") +
    theme()+
  labs(
    x = "Days",
    fill = "Trt"
  )
```